---
title: "Enhanced Time Series Analysis with Change Point Detection"
author: "India Climate Study - Temporal Analysis"
date: "`r Sys.Date()`"
format:
  html:
    theme: cosmo
    toc: true
    code-fold: false
---

# Enhanced Time Series Preparation and Analysis

This document implements advanced time series analysis including change point detection, trend analysis, and comprehensive temporal pattern identification for India's climate data.

## Setup

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load configuration and data
config <- readRDS("data/enhanced_config.rds")
india_dataset <- readRDS("data/processed/complete_india_dataset.rds")
performance_monitor <- readRDS("data/processed/performance_monitor.rds")

# Load required packages
library(terra)
library(ggplot2)
library(dplyr)
library(lubridate)
library(forecast)
library(tseries)
library(changepoint)
library(cowplot)
library(viridis)
library(ggridges)
library(tidyr)
```

## Create Synthetic Time Series from Spatial Data

```{r create-timeseries}
# Create comprehensive time series from spatial climate data
create_india_timeseries <- function(spatial_data, start_year = 2000, end_year = 2023) {
  
  cat("Creating synthetic time series from spatial data...\n")
  
  # Create date sequence
  dates <- seq(as.Date(paste0(start_year, "-01-01")), 
               as.Date(paste0(end_year, "-12-31")), 
               by = "month")
  
  n_months <- length(dates)
  n_years <- end_year - start_year + 1
  
  # Function to extract spatial mean and add temporal variation
  extract_timeseries <- function(spatial_raster, variable_name) {
    
    if(is.null(spatial_raster)) return(NULL)
    
    # Get spatial mean for each month/layer
    if(nlyr(spatial_raster) == 12) {
      # Monthly climatology data
      monthly_means <- sapply(1:12, function(i) {
        global(spatial_raster[[i]], fun = "mean", na.rm = TRUE)[[1]]
      })
      
      # Repeat pattern for each year and add variation
      base_series <- rep(monthly_means, n_years)
      
    } else {
      # Annual data - create monthly interpolation
      annual_mean <- global(spatial_raster, fun = "mean", na.rm = TRUE)[[1]]
      
      # Create synthetic seasonal pattern
      if(grepl("temp", variable_name, ignore.case = TRUE)) {
        # Temperature seasonal pattern for India
        seasonal_pattern <- 2 * sin(2 * pi * (1:12) / 12 - pi/2) # Peak in summer
      } else {
        # Precipitation seasonal pattern (monsoon)
        monsoon_pattern <- c(0.2, 0.3, 0.4, 0.6, 1.5, 4.0, 6.0, 5.0, 3.0, 1.0, 0.4, 0.2)
        seasonal_pattern <- monsoon_pattern / mean(monsoon_pattern)
      }
      
      monthly_pattern <- annual_mean * seasonal_pattern
      base_series <- rep(monthly_pattern, n_years)
    }
    
    # Add realistic temporal variation
    set.seed(config$random_seed + match(variable_name, names(spatial_data)))
    
    # Trend component
    if(grepl("temp", variable_name, ignore.case = TRUE)) {
      # Warming trend
      trend <- seq(0, 1.5, length.out = n_months)  # 1.5°C warming over period
    } else {
      # Precipitation trend (slight increase)
      trend <- seq(0, 0.1, length.out = n_months)  # 10% increase
      trend <- trend * base_series  # Multiplicative for precipitation
    }
    
    # Interannual variability
    if(grepl("temp", variable_name, ignore.case = TRUE)) {
      noise_sd <- 0.8  # Temperature variability
    } else {
      noise_sd <- base_series * 0.3  # 30% precipitation variability
    }
    
    # ENSO-like cycles (3-7 year periods)
    enso_cycle <- 0.5 * sin(2 * pi * seq_along(dates) / (4.5 * 12))
    
    # Create final time series
    time_series <- base_series + trend + enso_cycle + rnorm(n_months, 0, noise_sd)
    
    # Ensure precipitation is non-negative
    if(grepl("prec", variable_name, ignore.case = TRUE)) {
      time_series <- pmax(time_series, 0)
    }
    
    return(time_series)
  }
  
  # Create time series for all variables
  ts_data <- data.frame(Date = dates)
  
  # Create default variables if none exist
  if(is.null(spatial_data) || length(spatial_data) == 0) {
    cat("No spatial data found, creating synthetic climate variables...\n")
    
    # Create synthetic temperature data (tavg)
    temp_base <- 25  # Base temperature for India
    seasonal_temp <- 3 * sin(2 * pi * (1:12) / 12 - pi/2)  # Seasonal variation
    temp_pattern <- rep(temp_base + seasonal_temp, n_years)
    temp_trend <- seq(0, 1.5, length.out = n_months)  # Warming trend
    temp_noise <- rnorm(n_months, 0, 0.8)
    ts_data$tavg <- temp_pattern + temp_trend + temp_noise
    
    # Create synthetic precipitation data (prec)
    monsoon_pattern <- c(20, 30, 40, 80, 150, 300, 400, 350, 200, 100, 40, 25)
    prec_pattern <- rep(monsoon_pattern, n_years)
    prec_trend <- seq(0, 0.1, length.out = n_months) * prec_pattern
    prec_noise <- rnorm(n_months, 0, prec_pattern * 0.2)
    ts_data$prec <- pmax(prec_pattern + prec_trend + prec_noise, 0)
    
  } else {
    # Process existing spatial data
    for(var_name in names(spatial_data)) {
      if(!is.null(spatial_data[[var_name]])) {
        ts_values <- extract_timeseries(spatial_data[[var_name]], var_name)
        if(!is.null(ts_values)) {
          ts_data[[var_name]] <- ts_values
        }
      }
    }
    
    # Rename temperature and precipitation columns explicitly
    for(col_name in names(ts_data)) {
      if(grepl("temp|tavg", col_name, ignore.case = TRUE) && col_name != "tavg") {
        names(ts_data)[names(ts_data) == col_name] <- "tavg"
      } else if(grepl("prec|precip", col_name, ignore.case = TRUE) && col_name != "prec") {
        names(ts_data)[names(ts_data) == col_name] <- "prec"
      }
    }
  }
  
  # Add temporal features
  ts_data <- ts_data %>%
    mutate(
      Year = year(Date),
      Month = month(Date),
      Season = case_when(
        Month %in% c(12, 1, 2) ~ "Winter",
        Month %in% c(3, 4, 5) ~ "Pre-Monsoon",
        Month %in% c(6, 7, 8, 9) ~ "Monsoon",
        Month %in% c(10, 11) ~ "Post-Monsoon"
      ),
      Decade = paste0(floor((Year - 2000) / 10) * 10 + 2000, "s"),
      MonthName = month.name[Month]
    )
  
  return(ts_data)
}

# Create comprehensive time series
india_timeseries <- create_india_timeseries(
  india_dataset$india_climate_data,
  start_year = 2000,
  end_year = 2023
)

cat("Time series created with", nrow(india_timeseries), "observations\n")
cat("Variables:", ncol(india_timeseries) - 6, "\n")  # Minus date/temporal columns
cat("Period:", min(india_timeseries$Date), "to", max(india_timeseries$Date), "\n")

# Debug: Check column names
cat("Available columns:", paste(colnames(india_timeseries), collapse = ", "), "\n")

performance_monitor$log_performance("timeseries_creation")
```

## Advanced Change Point Detection

```{r changepoint-analysis}
# Comprehensive change point detection using changepoint package
detect_climate_changepoints <- function(timeseries_data, variables = c("tavg", "prec")) {
  
  changepoint_results <- list()
  
  # Filter variables to only those that exist
  available_vars <- intersect(variables, names(timeseries_data))
  
  if(length(available_vars) == 0) {
    cat("No climate variables found for change point analysis\n")
    return(list())
  }
  
  for(var_name in available_vars) {
    cat(paste("Analyzing change points for", var_name, "...\n"))
    
    ts_values <- timeseries_data[[var_name]]
    ts_values <- ts_values[!is.na(ts_values)]
    
    if(length(ts_values) < 50) {
      cat(paste("Insufficient data for", var_name, "\n"))
      next
    }
    
    # Multiple change point detection methods
    cp_results <- list()
    
    # Method 1: PELT (Pruned Exact Linear Time) - Changes in mean
    tryCatch({
      cp_pelt_mean <- cpt.mean(ts_values, method = "PELT", minseglen = 12)  # Min 1 year
      cp_results$pelt_mean <- list(
        method = "PELT_Mean",
        changepoints = cpts(cp_pelt_mean),
        dates = if(length(cpts(cp_pelt_mean)) > 0) timeseries_data$Date[cpts(cp_pelt_mean)] else as.Date(character(0)),
        n_changepoints = ncpts(cp_pelt_mean)
      )
    }, error = function(e) {
      cat(paste("PELT (mean) failed for", var_name, ":", e$message, "\n"))
    })
    
    # Method 2: PELT - Changes in variance
    tryCatch({
      cp_pelt_var <- cpt.var(ts_values, method = "PELT", minseglen = 12)
      cp_results$pelt_var <- list(
        method = "PELT_Variance",
        changepoints = cpts(cp_pelt_var),
        dates = if(length(cpts(cp_pelt_var)) > 0) timeseries_data$Date[cpts(cp_pelt_var)] else as.Date(character(0)),
        n_changepoints = ncpts(cp_pelt_var)
      )
    }, error = function(e) {
      cat(paste("PELT (variance) failed for", var_name, ":", e$message, "\n"))
    })
    
    # Method 3: Binary Segmentation - Changes in mean
    tryCatch({
      cp_binseg <- cpt.mean(ts_values, method = "BinSeg", Q = 5, minseglen = 12)
      cp_results$binseg <- list(
        method = "BinSeg",
        changepoints = cpts(cp_binseg),
        dates = if(length(cpts(cp_binseg)) > 0) timeseries_data$Date[cpts(cp_binseg)] else as.Date(character(0)),
        n_changepoints = ncpts(cp_binseg)
      )
    }, error = function(e) {
      cat(paste("BinSeg failed for", var_name, ":", e$message, "\n"))
    })
    
    # Method 4: At Most One Change (AMOC)
    tryCatch({
      cp_amoc <- cpt.mean(ts_values, method = "AMOC", minseglen = 24)  # Min 2 years
      cp_results$amoc <- list(
        method = "AMOC",
        changepoints = cpts(cp_amoc),
        dates = if(length(cpts(cp_amoc)) > 0) timeseries_data$Date[cpts(cp_amoc)] else as.Date(character(0)),
        n_changepoints = ncpts(cp_amoc)
      )
    }, error = function(e) {
      cat(paste("AMOC failed for", var_name, ":", e$message, "\n"))
    })
    
    # Method 5: PELT for mean and variance combined
    tryCatch({
      cp_pelt_meanvar <- cpt.meanvar(ts_values, method = "PELT", minseglen = 12)
      cp_results$pelt_meanvar <- list(
        method = "PELT_MeanVar",
        changepoints = cpts(cp_pelt_meanvar),
        dates = if(length(cpts(cp_pelt_meanvar)) > 0) timeseries_data$Date[cpts(cp_pelt_meanvar)] else as.Date(character(0)),
        n_changepoints = ncpts(cp_pelt_meanvar)
      )
    }, error = function(e) {
      cat(paste("PELT (meanvar) failed for", var_name, ":", e$message, "\n"))
    })
    
    changepoint_results[[var_name]] <- cp_results
    
    cat(paste("Change point analysis completed for", var_name, "with", length(cp_results), "methods\n"))
  }
  
  return(changepoint_results)
}

# Detect change points
changepoint_results <- detect_climate_changepoints(
  india_timeseries, 
  variables = c("tavg", "prec")
)

# Summarize change point results
summarize_changepoints <- function(cp_results) {
  summary_list <- list()
  
  for(var_name in names(cp_results)) {
    var_summary <- data.frame(
      Variable = character(),
      Method = character(),
      N_Changepoints = integer(),
      Key_Dates = character(),
      stringsAsFactors = FALSE
    )
    
    for(method in names(cp_results[[var_name]])) {
      cp_data <- cp_results[[var_name]][[method]]
      key_dates <- if(length(cp_data$dates) > 0) {
        paste(head(as.character(cp_data$dates), 3), collapse = ", ")
      } else {
        "None"
      }
      
      var_summary <- rbind(var_summary, data.frame(
        Variable = var_name,
        Method = cp_data$method,
        N_Changepoints = cp_data$n_changepoints,
        Key_Dates = key_dates,
        stringsAsFactors = FALSE
      ))
    }
    
    summary_list[[var_name]] <- var_summary
  }
  
  return(summary_list)
}

cp_summary <- summarize_changepoints(changepoint_results)

cat("\n=== CHANGE POINT DETECTION SUMMARY ===\n")
for(var in names(cp_summary)) {
  cat(paste("Variable:", var, "\n"))
  if(nrow(cp_summary[[var]]) > 0) {
    print(cp_summary[[var]])
  } else {
    cat("No change points detected\n")
  }
  cat("\n")
}

performance_monitor$log_performance("changepoint_analysis")
```

## Trend Analysis and Decomposition

```{r trend-analysis}
# Advanced trend analysis
analyze_climate_trends <- function(timeseries_data, variables = c("tavg", "prec")) {
  
  trend_results <- list()
  
  # Filter variables to only those that exist
  available_vars <- intersect(variables, names(timeseries_data))
  
  if(length(available_vars) == 0) {
    cat("No climate variables found for trend analysis\n")
    return(list())
  }
  
  for(var_name in available_vars) {
    cat(paste("Analyzing trends for", var_name, "...\n"))
    
    # Prepare time series object
    ts_data <- ts(timeseries_data[[var_name]], 
                  start = c(2000, 1), 
                  frequency = 12)
    
    # Time series decomposition
    decomp <- tryCatch({
      decompose(ts_data, type = "additive")
    }, error = function(e) {
      cat(paste("Decomposition failed for", var_name, "\n"))
      return(NULL)
    })
    
    # Linear trend analysis
    time_index <- 1:length(ts_data)
    trend_model <- lm(as.numeric(ts_data) ~ time_index)
    trend_slope <- coef(trend_model)[2]
    trend_pvalue <- summary(trend_model)$coefficients[2, 4]
    
    # Mann-Kendall trend test
    mk_test <- tryCatch({
      mk_result <- trend.test(as.numeric(ts_data), alternative = "two.sided")
      list(
        statistic = mk_result$statistic,
        p_value = mk_result$p.value,
        trend_direction = if(mk_result$statistic > 0) "Increasing" else "Decreasing"
      )
    }, error = function(e) {
      list(statistic = NA, p_value = NA, trend_direction = "Unknown")
    })
    
    # Seasonal trend analysis
    seasonal_trends <- timeseries_data %>%
      filter(!is.na(.data[[var_name]])) %>%
      group_by(Month) %>%
      summarise(
        monthly_mean = mean(.data[[var_name]], na.rm = TRUE),
        monthly_trend = {
          if(length(unique(Year)) > 1) {
            lm_model <- lm(.data[[var_name]] ~ Year)
            coef(lm_model)[2]
          } else {
            0
          }
        },
        monthly_trend_pval = {
          if(length(unique(Year)) > 1) {
            lm_model <- lm(.data[[var_name]] ~ Year)
            summary(lm_model)$coefficients[2, 4]
          } else {
            1
          }
        },
        .groups = 'drop'
      ) %>%
      mutate(
        month_name = month.name[Month],
        trend_significant = monthly_trend_pval < 0.05
      )
    
    # Compile results
    trend_results[[var_name]] <- list(
      decomposition = decomp,
      linear_trend = list(
        slope_per_month = trend_slope,
        slope_per_decade = trend_slope * 120,  # 10 years * 12 months
        p_value = trend_pvalue,
        significant = trend_pvalue < 0.05
      ),
      mann_kendall = mk_test,
      seasonal_trends = seasonal_trends,
      summary = list(
        variable = var_name,
        overall_trend_direction = if(trend_slope > 0) "Increasing" else "Decreasing",
        trend_magnitude = abs(trend_slope * 120),
        statistical_significance = if(trend_pvalue < 0.05) "Significant" else "Not Significant"
      )
    )
    
    cat(paste("Trend analysis completed for", var_name, "\n"))
  }
  
  return(trend_results)
}

# Analyze trends
trend_results <- analyze_climate_trends(india_timeseries, c("tavg", "prec"))

# Display trend summaries
cat("\n=== CLIMATE TREND ANALYSIS SUMMARY ===\n")
for(var in names(trend_results)) {
  summary_info <- trend_results[[var]]$summary
  linear_info <- trend_results[[var]]$linear_trend
  
  cat(paste("Variable:", summary_info$variable, "\n"))
  cat(paste("Trend Direction:", summary_info$overall_trend_direction, "\n"))
  cat(paste("Magnitude (per decade):", round(summary_info$trend_magnitude, 4), "\n"))
  cat(paste("Statistical Significance:", summary_info$statistical_significance, "\n"))
  cat(paste("P-value:", format.pval(linear_info$p_value), "\n"))
  
  # Mann-Kendall results
  mk_info <- trend_results[[var]]$mann_kendall
  if(!is.na(mk_info$p_value)) {
    cat(paste("Mann-Kendall trend:", mk_info$trend_direction, 
              "(p-value:", format.pval(mk_info$p_value), ")\n"))
  }
  cat("\n")
}

performance_monitor$log_performance("trend_analysis")
```

## Enhanced Visualization Dashboard

```{r enhanced-visualization, fig.width=16, fig.height=14}
# Create comprehensive temporal analysis dashboard
create_temporal_dashboard <- function(timeseries_data, trend_results, changepoint_results) {
  
  plots <- list()
  
  # Get available climate variables dynamically
  climate_vars <- intersect(c("tavg", "prec"), names(timeseries_data))
  
  if(length(climate_vars) == 0) {
    cat("No climate variables found for visualization\n")
    return(list())
  }
  
  # Plot 1: Time series with trends and change points
  for(var in climate_vars) {
    # Prepare plot data
    plot_data <- timeseries_data %>%
      select(Date, all_of(var)) %>%
      filter(!is.na(.data[[var]]))
    
    # Base time series plot
    ts_plot <- ggplot(plot_data, aes(x = Date, y = .data[[var]])) +
      geom_line(alpha = 0.7, color = "steelblue") +
      geom_smooth(method = "lm", se = TRUE, color = "red", alpha = 0.3) +
      labs(title = paste("Time Series Analysis:", var),
           x = "Date", y = if(var == "tavg") "Temperature (°C)" else "Precipitation (mm)") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
    # Add change points if available
    if(var %in% names(changepoint_results)) {
      for(method in names(changepoint_results[[var]])) {
        cp_data <- changepoint_results[[var]][[method]]
        if(length(cp_data$dates) > 0) {
          ts_plot <- ts_plot +
            geom_vline(xintercept = as.numeric(cp_data$dates), 
                      color = "orange", linetype = "dashed", alpha = 0.7)
        }
      }
    }
    
    plots[[paste0(var, "_timeseries")]] <- ts_plot
  }
  
  # Plot 2: Seasonal patterns (only if both variables exist)
  if(all(c("tavg", "prec") %in% climate_vars)) {
    seasonal_plot <- timeseries_data %>%
      select(Date, Month, Season, tavg, prec) %>%
      filter(!is.na(tavg), !is.na(prec)) %>%
      pivot_longer(cols = c(tavg, prec), names_to = "Variable", values_to = "Value") %>%
      mutate(
        Variable = case_when(
          Variable == "tavg" ~ "Temperature (°C)",
          Variable == "prec" ~ "Precipitation (mm)",
          TRUE ~ Variable
        )
      ) %>%
      ggplot(aes(x = factor(Month), y = Value, fill = Season)) +
      geom_boxplot(alpha = 0.7) +
      facet_wrap(~Variable, scales = "free_y", ncol = 1) +
      scale_fill_viridis_d(name = "Season") +
      labs(title = "Seasonal Climate Patterns",
           x = "Month", y = "Value") +
      theme_minimal() +
      theme(legend.position = "bottom")
    
    plots$seasonal <- seasonal_plot
  }
  
  # Plot 3: Interannual variability (only if both variables exist)
  if(all(c("tavg", "prec") %in% climate_vars)) {
    annual_data <- timeseries_data %>%
      group_by(Year) %>%
      summarise(
        Annual_Temp = mean(tavg, na.rm = TRUE),
        Annual_Precip = sum(prec, na.rm = TRUE),
        .groups = 'drop'
      ) %>%
      pivot_longer(cols = c(Annual_Temp, Annual_Precip), 
                   names_to = "Variable", values_to = "Value")
    
    annual_plot <- ggplot(annual_data, aes(x = Year, y = Value, color = Variable)) +
      geom_line(size = 1) +
      geom_smooth(method = "loess", se = TRUE, alpha = 0.3) +
      facet_wrap(~Variable, scales = "free_y", ncol = 1) +
      scale_color_manual(values = c("Annual_Temp" = "red", "Annual_Precip" = "blue")) +
      labs(title = "Interannual Climate Variability",
           x = "Year", y = "Value") +
      theme_minimal() +
      theme(legend.position = "none")
    
    plots$annual <- annual_plot
  }
  
  # Plot 4: Trend decomposition (for temperature if available)
  if("tavg" %in% names(trend_results) && !is.null(trend_results$tavg$decomposition)) {
    decomp_data <- trend_results$tavg$decomposition
    
    decomp_df <- data.frame(
      Date = timeseries_data$Date,
      Observed = as.numeric(decomp_data$x),
      Trend = as.numeric(decomp_data$trend),
      Seasonal = as.numeric(decomp_data$seasonal),
      Residual = as.numeric(decomp_data$random)
    ) %>%
      pivot_longer(cols = -Date, names_to = "Component", values_to = "Value")
    
    decomp_plot <- ggplot(decomp_df, aes(x = Date, y = Value)) +
      geom_line(color = "steelblue", alpha = 0.8) +
      facet_wrap(~Component, scales = "free_y", ncol = 1) +
      labs(title = "Temperature Time Series Decomposition",
           x = "Date", y = "Value") +
      theme_minimal() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
    plots$decomposition <- decomp_plot
  }
  
  return(plots)
}

# Generate temporal dashboard
temporal_plots <- create_temporal_dashboard(india_timeseries, trend_results, changepoint_results)

# Display plots in organized layout
if(length(temporal_plots) >= 4) {
  # Create multi-page dashboard
  available_ts_plots <- temporal_plots[grepl("_timeseries$", names(temporal_plots))]
  
  if(length(available_ts_plots) >= 2) {
    page1 <- cowplot::plot_grid(
      plotlist = available_ts_plots[1:2],
      ncol = 1
    )
  } else if(length(available_ts_plots) == 1) {
    page1 <- available_ts_plots[[1]]
  }
  
  other_plots <- temporal_plots[!grepl("_timeseries$", names(temporal_plots))]
  if(length(other_plots) >= 2) {
    page2 <- cowplot::plot_grid(
      plotlist = other_plots[1:2],
      ncol = 2
    )
  } else if(length(other_plots) == 1) {
    page2 <- other_plots[[1]]
  }
  
  if(exists("page1")) print(page1)
  if(exists("page2")) print(page2)
  
  # Save plots
  if(exists("page1")) {
    ggsave(file.path(config$output_dir, "plots", "temporal_analysis_page1.png"), 
           page1, width = 16, height = 10, dpi = 300, bg = "white")
  }
  
  if(exists("page2")) {
    ggsave(file.path(config$output_dir, "plots", "temporal_analysis_page2.png"), 
           page2, width = 16, height = 8, dpi = 300, bg = "white")
  }
  
  if("decomposition" %in% names(temporal_plots)) {
    print(temporal_plots$decomposition)
    ggsave(file.path(config$output_dir, "plots", "temperature_decomposition.png"), 
           temporal_plots$decomposition, width = 14, height = 10, dpi = 300, bg = "white")
  }
  
} else {
  # Display available plots
  for(plot_name in names(temporal_plots)) {
    print(temporal_plots[[plot_name]])
  }
}

performance_monitor$log_performance("visualization_complete")
```

## Save Enhanced Time Series Data

```{r save-timeseries-data}
# Compile comprehensive time series analysis results
enhanced_timeseries_results <- list(
  
  # Time series data
  india_timeseries = india_timeseries,
  
  # Analysis results
  changepoint_results = changepoint_results,
  trend_results = trend_results,
  changepoint_summary = cp_summary,
  
  # Metadata
  analysis_metadata = list(
    timestamp = Sys.time(),
    time_period = paste(min(india_timeseries$Date), "to", max(india_timeseries$Date)),
    variables_analyzed = intersect(c("tavg", "prec"), names(india_timeseries)),
    changepoint_methods = c("PELT_Mean", "PELT_Variance", "PELT_MeanVar", "BinSeg", "AMOC"),
    trend_methods = c("Linear Regression", "Mann-Kendall"),
    total_observations = nrow(india_timeseries),
    analysis_frequency = "Monthly"
  )
)

# Save individual components
saveRDS(india_timeseries, "data/processed/india_climate_timeseries.rds")
saveRDS(changepoint_results, "data/processed/changepoint_analysis.rds")
saveRDS(trend_results, "data/processed/trend_analysis.rds")
saveRDS(enhanced_timeseries_results, "data/processed/complete_timeseries_analysis.rds")

# Export CSV data
write.csv(india_timeseries, "data/processed/india_climate_timeseries.csv", row.names = FALSE)

# Export analysis summaries
if(length(trend_results) > 0) {
  trend_summary_df <- do.call(rbind, lapply(names(trend_results), function(var) {
    summary_info <- trend_results[[var]]$summary
    linear_info <- trend_results[[var]]$linear_trend
    mk_info <- trend_results[[var]]$mann_kendall
    
    data.frame(
      Variable = var,
      Trend_Direction = summary_info$overall_trend_direction,
      Slope_Per_Decade = round(linear_info$slope_per_decade, 6),
      Linear_P_Value = linear_info$p_value,
      Linear_Significant = linear_info$significant,
      MK_Trend = mk_info$trend_direction,
      MK_P_Value = mk_info$p_value,
      stringsAsFactors = FALSE
    )
  }))
  
  write.csv(trend_summary_df, "data/processed/climate_trend_summary.csv", row.names = FALSE)
}

# Export change point summary
if(length(cp_summary) > 0) {
  cp_summary_df <- do.call(rbind, cp_summary)
  write.csv(cp_summary_df, "data/processed/changepoint_summary.csv", row.names = FALSE)
}

performance_monitor$log_performance("timeseries_saving")

cat("Enhanced time series analysis results saved\n")
```

## Summary

```{r final-summary}
cat("\n=== ENHANCED TIME SERIES ANALYSIS COMPLETE ===\n")
cat("Time series period:", paste(min(india_timeseries$Date), "to", max(india_timeseries$Date)), "\n")
cat("Total observations:", nrow(india_timeseries), "\n")
cat("Variables with time series:", length(intersect(c("tavg", "prec"), names(india_timeseries))), "\n")

# Summary of key findings
if("tavg" %in% names(trend_results)) {
  temp_trend <- trend_results$tavg$summary
  cat("Temperature trend:", temp_trend$overall_trend_direction, 
      "(", round(temp_trend$trend_magnitude, 4), "per decade,", 
      temp_trend$statistical_significance, ")\n")
}

if("prec" %in% names(trend_results)) {
  precip_trend <- trend_results$prec$summary
  cat("Precipitation trend:", precip_trend$overall_trend_direction,
      "(", round(precip_trend$trend_magnitude, 4), "per decade,",
      precip_trend$statistical_significance, ")\n")
}

# Change point summary
total_changepoints <- sum(sapply(cp_summary, function(x) sum(x$N_Changepoints, na.rm = TRUE)))
cat("Total change points detected:", total_changepoints, "\n")

cat("\nFiles Created:\n")
cat("- data/processed/india_climate_timeseries.rds\n")
cat("- data/processed/changepoint_analysis.rds\n")
cat("- data/processed/trend_analysis.rds\n")
cat("- data/processed/complete_timeseries_analysis.rds\n")
cat("- data/processed/india_climate_timeseries.csv\n")

if(length(trend_results) > 0) {
  cat("- data/processed/climate_trend_summary.csv\n")
}

if(file.exists("data/processed/changepoint_summary.csv")) {
  cat("- data/processed/changepoint_summary.csv\n")
}

# List visualization files
viz_files <- list.files(file.path(config$output_dir, "plots"), 
                       pattern = "temporal|decomposition", full.names = FALSE)
if(length(viz_files) > 0) {
  cat("Visualization files:\n")
  for(file in viz_files) {
    cat(paste("- outputs/plots/", file, "\n", sep = ""))
  }
}

cat("\nNext Step: Run 05-enhanced-arima-modeling.qmd\n")
```
